---
title: "Final Project Prelimiary EDA"
author: "Xunyuan Sun & Jing Du"
output: html_notebook
---

## Guiding Question
```{r}
library(DataComputing)
library(mosaic)
```


## Become acquainted with the data sources

1. Where did you find them?
We found these data sources from kaggle.

2. Who collected/maintains them?
The user name is MuonNeutrino

3. When & Why were they originally collected?
These data sources was updated one years ago.

4. What does a case represent in each data source, and how many total cases are available?
In acs2015_census_tract_data, each case for 2015 means each census tract in the U.S.
In acs2017_census_tract_data, each case for 2017 means each census tract in the U.S.

5. What are some of the variables that you plan to use
TractId, State, County, TotalPop, Men, Women, Hispanic, White, Black, Native, Asian, Pacific, Income

```{r}
Census_Tract_2015 <-
  read.csv("https://raw.githubusercontent.com/XYSunPSU/Final-Project/master/acs2015_census_tract_data.csv", header = T)

Census_Tract_2017 <-
  read.csv("https://raw.githubusercontent.com/XYSunPSU/Final-Project/master/acs2017_census_tract_data.csv", header = T)
```


```{r}
Census_Tract_2015
```


```{r}
Census_Tract_2017
```

## Explore intuition related to research question
1. Create some informative plots and summary statistics 
```{r}
graphdata <-
  Census_Tract_2015%>%
  filter(State == "Pennsylvania")

ggplot(data=graphdata,aes(x=County,y=TotalPop ))+geom_bar(stat='identity',position='stack', width=.9) 
```
```{r}

graphdata <-
  Census_Tract_2015%>%
  filter(State == "Alabama")

ggplot(data=graphdata,aes(x=County,y=TotalPop ))+geom_bar(stat='identity',position='stack', width=.9) 

```
```{r}

incomedata <-
  Census_Tract_2015%>%
  filter(State == "Pennsylvania")

ggplot(data=incomedata,aes(x=County,y=Income ))+geom_bar(stat='identity',position='stack', width=.9) 

```
```{r}
incomedata2 <-
  Census_Tract_2015%>%
  filter(State == "Pennsylvania")

ggplot(data=incomedata2,aes(x=County,y=Income ))+geom_bar(stat='identity',position='stack', width=.9) 
```

  

2. Describe preliminary observations and intuition about the research question

## Two or more data sources


1.  acs2017_census_tract_data.csv

2.  acs2015_census_tract_data.csv

